
%I have many people to thank who have advised and supported me, and without them I could not have written this thesis. I am very much a product of the people around me, and every one of my friends, colleagues, and academic advisors have shaped me as a scientist and I cannot thank them enough.

First and foremost, I would like to thank Prof. Tom Milne and Prof. Marella de Bruijn for being fantastic and supportive mentors. Working in Marella's lab gave me my first taste of academia, and helped me develop the critical groundwork for being an academic scientist. Tom has been an incredible guide in navigating my DPhil project, and has always pushed me to be imaginative and driven in my work. 
\vspace{-8pt}

The Milne lab has been home to many lovely colleagues, and I am grateful to work alongside all of them. Nick Crump has taught me many techniques and pushed my critical thinking. I will always remember the late-night lab chats, and spurious analysis sessions. I-Jun Lau has been kind and supportive, and always ready to talk science. To Nick and I-Jun: I'll never stop singing "do you like tdTomato". Alastair Smith has been good to talk to (and complain to) about coding, though I won't stop defending R over Python. Vassi Sharlandjieva, Nicole Jackson and Catherine Chahrour have been bright personalities in the lab, and have brought a little life back into work-life balance. Ross Thorne left the lab before I joined, but his initial work on networks was critical to my thesis, and his viva presentation was immensely inspirational. Outside the lab, Siobhan Rice and Simona Valletta were a fantastic help with mouse experiments.
\vspace{-8pt}

In the de Bruijn lab, Vincent Frontera was always entertaining to work with, and taught me a lot about flow cytometry. I'm thankful for Christina Rode, who patiently taught me many basic lab techniques, and was always a positive personality. Dominic Owens was a great colleague, who helped me with his CRISPR and capture-C knowledge, and always had a great story to tell. Emanuele Azzoni taught me how to perform whole mount imaging, which are not only scientifically useful, but are artistically beautiful. Lucas Greder performed experiments that have been the seed for my network analyses. Giorgio Anselmi and Alessandro Cavallo have always been ready and willing to talk science, particularly single cell analyses, and have offered great insight. Matt Nicholls has been meticulous in organising the lab, and I am grateful for the times he has kept me company in the mouse facility. Naeema Mehmood has been lovely to talk to - and I appreciate being able to talk about the latest Marvel films!
\vspace{-8pt}

Most importantly, I would like to thank my family. My mum, Angela, pushed me from a young age to do well academically. I'm sure I can only do maths because we recited the times table so often! My dad, Ritchie, has always supported me, and I inherited my love of fancy tech from him. My sister (the "cooler" sibling), Rachael, has been consistently supportive.% Finally, Claire has been incredibly supportive and patient while I've written this thesis!