\textit{Joint first authors indicated with *}

\noindent
\textbf{Harman JR}*, Thorne R*, Jamilly M, Tapia M, Crump NT, Rice S, Beveridge R, Morrissey E, de Bruijn MFTR, Roberts I, Roy A, Fulga TA, Milne TA. A KMT2A-AFF1 gene regulatory network highlights the role of core transcription factors and reveals the regulatory logic of key downstream target genes. Genome Res. 2021 Jun 4;31(7):1159–73

\noindent
\textbf{Harman JR}*, Greder L*, Swiers G, Owens DDG, Suciu M, Downes D, Telenius J, Frontera V, Antoniou S, Repapi E, Azzoni E, de Bruijn MFTR. A gene regulatory network of the endothelial-to-hematopoietic transition reveals a Runx1:Ikzf1:Notch feed-forward loop. \textit{Manuscript in preparation}.

\noindent
Fidyt K, Pastorczak A, Cyran J, Crump NT, Goral A, Madzio J, Muchowicz A, Poprzeczko M, Domka K, Komorowski L, Winiarska M, \textbf{Harman JR}, Siudakowska K, Graczyk-Jarzynka A, Patkowska E, Lech-Maranda E, Mlynarski W, Golab J, Milne TA, Firczuk M. Potent, p53-independent induction of NOXA sensitizes MLL-rearranged B-cell acute lymphoblastic leukemia cells to venetoclax. Oncogene. 2022 Mar;41(11):1600-1609

\noindent
Owens DDG, Anselmi G, Oudelaar AM, Downes DJ, Cavallo A, \textbf{Harman JR}, Schwessinger R, Bucakci A, Greder L, de Ornellas S, Jeziorska D, Telenius J, Hughes JR, de Bruijn MFTR. Dynamic Runx1 chromatin boundaries affect gene expression in hematopoietic development. Nat Commun. 2022 Feb 9;13(1):773

\noindent
Azzoni E, Frontera V, Anselmi G, Rode C, James C, Deltcheva EM, Demian AS, Brown J, Barone C, Patelli A, \textbf{Harman JR}, Nicholls M, Conway SJ, Morrissey E, Jacobsen SEW, Sparrow DB, Harris AL, Enver T, de Bruijn MFTR. The onset of circulation triggers a metabolic switch required for endothelial to hematopoietic transition. Cell Rep. 2021 Dec 14;37(11):110103

\noindent
Godfrey L, Crump NT, O'Byrne S, Lau IJ, Rice S, \textbf{Harman JR}, Jackson T, Elliott N, Buck G, Connor C, Thorne R, Knapp DJHF, Heidenreich O, Vyas P, Menendez P, Inglott S, Ancliff P, Geng H, Roberts I, Roy A, Milne TA. H3K79me2/3 controls enhancer-promoter interactions and activation of the pan-cancer stem cell marker PROM1/CD133 in MLL-AF4 leukemia cells. Leukemia. 2021 Jan;35(1):90-106

\noindent
Owens DDG, Caulder A, Frontera V, \textbf{Harman JR}, Allan AJ, Bucakci A, Greder L, Codner GF, Hublitz P, McHugh PJ, Teboul L, de Bruijn MFTR. Microhomologies are prevalent at Cas9-induced larger deletions. Nucleic Acids Res. 2019 Aug 22;47(14):7402-7417

\noindent
Godfrey L, Crump NT, Thorne R, Lau IJ, Repapi E, Dimou D, Smith AL, \textbf{Harman JR}, Telenius JM, Oudelaar A, Downes DJ, Vyas P, Hughes JR, Milne TA. DOT1L inhibition reveals a distinct subset of enhancers dependent on H3K79 methylation. Nature Communications. 10, 2803 (2019).

\noindent
\textbf{Conference attendance}

\noindent
Presentation: “Reconstruction of an EHT gene regulatory network identifies a Runx1:Ikzf1:Notch feed-forward loop”, International Society for Experimental Hematology. Online, August 2021.

\noindent
Poster: “RUNX1 gene regulatory network in leukaemia”, International RUNX conference. Seoul National University, Seoul, Korea. June 2019.
